---
title: "Ghana Data"
author: "Tomas Morley"
date: "5 July 2016"
output: pdf_document
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
# get packages
library(ggplot2)
library(dplyr)
library(stargazer)
# library(corrplot)
library(car)
library(rgl)
library(grid)
library(gridExtra)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# read in the data
filePath <- "C:/Users/Tomas/Documents/LEI"
GHA2010 <- readRDS(file.path(filePath, "data/GHA/GHA2010.rds"))
```

```{r echo=FALSE}
# read in functions
source(file.path(filePath, "functions/summTabs.R"))
source(file.path(filePath, "functions/removeOutliers.R"))
```

# Introduction 

This pdf document introduces and describes the Ghana data from the . . . . Yale survey. Below you will find exploratory analysis and a first basic analysis of the 2010 cross section. Data is available for one year 

## maps of Ghana

```{r echo=FALSE, message=FALSE, eval=FALSE}
# source in file the has the maps of Ghana
library(gridExtra)
source(file.path(filePath, "pro-gap/GHA/GHAmaps.R"))
grid.arrange(gmap1, gmap3, ncol=2)
```

```{r echo=FALSE}
# basic counts of available data
households <- group_by(GHA2010, reg) %>% summarise(households=length(unique(hhno)))
plots <- group_by(GHA2010, reg) %>% summarise("maize plots"=n())
fert <- group_by(GHA2010, reg) %>% summarise("plots with fert"=sum(N>0))
out <- left_join(households, plots) %>% left_join(fert)
knitr::kable(out)
```

# Univariate Analysis

Table 1 below shows summary statistics for key variables in the GHA 2010 data.

```{r echo=FALSE}
probs <- c(0.1, 0.25, 0.5, 0.75, 0.9)
varNames <- c("yld", "N", "P", "K", "herb", "fung", "pest",
              "lab", "asset", "area", "area_tot", "irrig",
              "maze_prc", "crop_count")
Nat <- summTab(GHA2010, varNames=varNames, probs=probs)
knitr::kable(Nat, caption="Summary Statistics for key variables")
```

Two variables key to this analysis are the nitrogen and yld variables. plotted below in level and log form figure ??.

```{r echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1, 2))
with(GHA2010, {
  hist(yld, breaks="FD", freq=FALSE, ylab="Density")
  lines(density(yld), lwd=2)
  lines(density(yld, adjust=0.5), lwd=1)
  rug(yld)
  box()
})
with(GHA2010, {
  hist(log(yld), breaks="FD", freq=FALSE, ylab="Density")
  lines(density(log(yld)), lwd=2)
  lines(density(log(yld), adjust=0.5), lwd=1)
  rug(log(yld))
  box()
})
```

And for nitrogen similar plots are shown in figure ??

```{r echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1, 2))
with(GHA2010, {
  hist(N[N>0 & N < 200], breaks="FD", freq=FALSE, ylab="Density")
  lines(density(N[N>0 & N < 200]), lwd=2)
  lines(density(N[N>0 & N < 200], adjust=0.5), lwd=1)
  rug(N[N>0 & N < 200])
  box()
})
with(GHA2010, {
  hist(log(N[N>0 & N < 200]), breaks="FD", freq=FALSE, ylab="Density")
  lines(density(log(N[N>0 & N < 200])), lwd=2)
  lines(density(log(N[N>0 & N < 200]), adjust=0.5), lwd=1)
  rug(log(N[N>0 & N < 200]))
  box()
})
```


# Multivariate Analysis

```{r eval=FALSE}
GHA2010_2 <-GHA2010[,varNames]
COR.DIS <- cor(GHA2010_2[complete.cases(GHA2010_2),])
corrplot(COR.DIS, method="square")
```

```{r }
GHA2010$lyld <- log(GHA2010$yld)
GHA2010$lN <- log(GHA2010$N + 1)
# GHA2010_2 <-GHA2010[,c(varNames, "lN", "lyld")]
# COR.DIS <- cor(GHA2010_2[complete.cases(GHA2010_2),])
# corrplot(COR.DIS, method="square")
```


```{r}
scatter3d(lyld ~ lN + crop_count, id.n=3, data=GHA2010)
```

```{r}
scatterplotMatrix(~lyld + lN + asset + lab, span=0.7, data=GHA2010)
```

# Building the Cobb Douglas model

```{r echo=FALSE, eval=FALSE}
# start with just a cobb douglass regression of
# yld against nitrogen
CD1 <- lm(log(yld) ~ log(I(N+1)), data = db1)
output <- round(summary(CD1)$coefficients, 3)
# row.names(output) <- c()
knitr::kable(output, caption="Cobb Douglas")
```

```{r echo=FALSE, eval=FALSE}
# Cobb Douglas regression - add in assets and labour
CD1 <- lm(log(yld) ~ log(I(N+1)) + log(asset + 1) + log(lab), data=db1)
output <- round(summary(CD1)$coefficients, 3)
# row.names(output) <- c()
knitr::kable(output, caption="Cobb Douglas")
```

```{r echo=FALSE, eval=FALSE}
CD1 <- lm(log(yld) ~ log(I(N+1)):REGNAME + log(asset + 1) + log(lab), data=db1)
output <- round(summary(CD1)$coefficients, 3)
# row.names(output) <- c()
knitr::kable(output, caption="Cobb Douglas")
```

```{r echo=FALSE, eval=FALSE}
CD1 <- lm(log(yld) ~ log(I(N+1)):REGNAME + log(asset + 1) + log(lab) + 
            #legume +
            herb + fung + log(area_tot) + irrig + crop_count + log(area),
          data=db1)
output <- round(summary(CD1)$coefficients, 3)
# row.names(output) <- c()
knitr::kable(output, caption="Cobb Douglas")
```


```{r echo=FALSE, eval=FALSE}
plot(CD1)

# look at residuals and outliers
rd <- residuals(CD1)
lev <- influence(CD1)$hat
rdstudent <- rstudent(CD1)

par(mfrow=c(2, 2))
plot(lev, main="Leverages")
plot(rd, main = "Residuals (rd)")
plot(rdstudent, main = "Studentized Residuals")

plot(residuals(CD1) ~ predict(CD1),
     xlab=expression(hat(eta)),ylab="rd", main="Residuals vs. fitted values (link scale)")
qqplot(residuals(CD1))
```

```{r echo=FALSE, eval=FALSE}
# Cobb Douglas regression - baseline region is Western Region
CD1 <- lm(log(yld) ~ log(I(N+1)) + log(asset + 1) + log(lab) + legume, data=GHA2010)
output <- round(summary(CD1)$coefficients, 3)
# row.names(output) <- c()
knitr::kable(output, caption="Cobb Douglas")
```

```{r echo=FALSE, eval=FALSE}
# Cobb Douglas regression - Nitrogen effect different across regions
CD1 <- lm(log(yld) ~ log(I(N+1)):reg + log(asset + 1) + log(lab), data=GHA2010)
output <- round(summary(CD1)$coefficients, 3)
# row.names(output) <- c()
knitr::kable(output, caption="Cobb Douglas")
```


```{r echo=FALSE, eval=FALSE}
# quadratic regression - basic regression
QUAD1 <- lm(yld ~ N + I(N^2), data=db1)
```

```{r echo=FALSE, eval=FALSE}
QUAD2 <- lm(yld ~ N:REGNAME + I(N^2):REGNAME, data=db1)
```

```{r echo=FALSE, eval=FALSE}
QUAD3 <- lm(yld ~ N:REGNAME + I(N^2):REGNAME + asset + I(asset^2), data=db1)
```

```{r echo=FALSE, eval=FALSE}
QUAD4 <- lm(yld ~ N:REGNAME + I(N^2):REGNAME + asset + I(asset^2) + area + I(area^2), data=db1)
```

```{r echo=FALSE, eval=FALSE}
QUAD5 <- lm(yld ~ N:REGNAME + I(N^2):REGNAME + asset + I(asset^2) + area + I(area^2) +
             manure + herb + fung +
              manure + mech + area_tot + I(area_tot^2) + irrig + lab + lab^2,
            data=db1)
```

```{r, results="asis", echo=FALSE, eval=FALSE}
stargazer(QUAD1, QUAD2, QUAD3, QUAD4, header=FALSE, type='text')
```
# Spatial Analysis