---
title: "Ghana yield gap analysis"
author: "Michiel van Dijk & Tom Morley"
date: "6 May, 2016"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#dataPath <- "C:/Users/Tomas/Documents/LEI/GHA_presentation"
dataPath <- "D:\\Dropbox\\Michiel_research\\2285000066 Africa Maize Yield Gap\\Analysis\\GHA\\Presentations\\GHA_presentation"
#dataPath <- "D:\\Data\\IPOP\\SurveyData\\GHA\\2010\\Data"
GHAdata <- readRDS(file.path(dataPath, "GHA.rds"))
library(ggplot2)
devtools::source_gist("33baa3a79c5cfef0f6df")
```

# EGC-ISSER Socioeconomic Panel Survey
- Collaboration between the Economic Growth Center (EGC) at Yale University and the  Institute of Statistical, Social, and Economic Research (ISSER) at the University of Ghana, Legon.
- Large-scale, nation-wide panel survey in Ghana that will extend for at least 15 years (3 year interval).
- First wave in 2010, 2013 wave in progress. 
- Regionally representative sample for the 10 regions of Ghana.
- 5009 households from 334 Enumeration Areas (EAs).


# Summary statistics for key variables
\footnotesize 
```{r, results='asis'}
GHAdata <- readRDS(file.path(dataPath, "GHA.rds"))
unwanted <- c("hhno", "fallowB", "fallowF", "REGION_CODE", "DISTRICT_CODE")
GHAdata2 <- GHAdata[!names(GHAdata) %in% unwanted]
suppressMessages(library(stargazer))
stargazer(as.data.frame(GHAdata2),  type = "latex", digits=2, header=FALSE, title = "Summary statistics")
```


# Data issue #1: Area measurement
```{r, echo=FALSE, message=FALSE}
GHAdata2 <- readRDS(file.path(dataPath, "GHAdata2.rds"))
ggplot(GHAdata2, aes(x=area)) + geom_histogram(binwidth=0.05, fill="red", color="black") +
  theme_bw() + ggtitle("Histogram of plot size")
```


# Data issue #2: Location is not reliable
```{r, echo=FALSE, message=FALSE}
library(haven)
Region_map <- readRDS(file.path(dataPath, "Region_map.rds"))
coords <- readRDS(file.path(dataPath, "Region_map_coords.rds"))

base <- ggplot(Region_map) + 
  geom_polygon(data=Region_map, aes(x=long, y=lat, group=group, fill=REGION_NAME), colour="black",
                size = 1) +
  scale_fill_manual(values = rainbow(n=12, alpha = .8)) +
  coord_map("mercator") +
  theme_map() +
  guides(fill=FALSE) +
  geom_text(data=coords, aes(label = REGION_NAME, x = x, y = y)) 

gps_data <- read_dta(file.path(dataPath, "data/ShiftedGPSData.dta"))
gps_plot <- base + geom_point(data=gps_data, aes(x=longitude, y = latitude), col="black", shape=16)
print(gps_plot)
```

# Data issue #2: Location is not reliable
- Code book with information on regions and districts does not match with data
- Data files seem to have impossible region-district combinations
- Best would be to use GPS coordinates but datafile is incomplete

# Data issue #3: EGC-ISSER survey versus case-study regions
```{r, warning=FALSE, echo=FALSE, message=FALSE}
library(maptools)
library(dplyr)
survey <- as.data.frame(do.call(rbind, getKMLcoordinates(file.path(dataPath, "VSA_V3_results.kml")))) %>% select(longitude = V1,latitude = V2)

gf2.2 <- readRDS(file.path(dataPath, "district_map.rds"))
dis_name = c("Savelugu Nanton", "Nkoranza")
dis_points <- readRDS(file.path(dataPath, "dis_points.rds"))


link <- ggplot(gf2.2, aes(long, lat)) +
  coord_map() +
  geom_polygon(aes(group=group),fill="white", colour="black") +
  guides(fill=FALSE) +
  geom_polygon(data=subset(gf2.2, NAME_2 %in% dis_name),
               aes(group=group), size=1, colour = "black", fill="light blue") +
  guides(fill=FALSE) +
  geom_text(data=dis_points, aes(label = DISTRICT_NAME, x = x, y = y)) +
  theme_map() + 
  geom_point(data=gps_data, aes(x=longitude, y = latitude), col="black", shape=16) +
  geom_point(data=survey, aes(x=longitude, y = latitude), col="red", shape=16)

print(link)
```

# Data issue #4: Fertilizer data
- Type of fertilizer (e.g DAP, UREA, etc.) is not presented (but might be recorded)
- Different units for fertilizer quantities:
    + Basket
    + Beer bottle
    + Maxi bag
    + Mini bag
    + Margarine tin
- But we miss converions factors...
