---
title: "Ghana Yield Gap Analysis"
author: "Michiel van Dijk & Tom Morley"
date: "10 April 2017"
output: beamer_presentation
root.dir: "C:/users/morle001/WEcR/GHAYG"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(pander)
```

```{r get_data, echo = FALSE, message = FALSE}
# Get workspace with all info
root <- "C:/users/morle001/WEcR/GHAYG"
source(file.path(root, "Code/GHA_fig.R"))
```

## Recommendations

- Largest yield gap is the technical yield gap, followed by the economic, allocative and technical efficiency yield gaps.

- Policies that close the allocative or economic yield gaps are very important. For example, subsidy programs which lower the cost of nitrogen will help to close the allocative yield gap.

- Determinants of inefficiency (still to come).

## Outline of analysis

- Working within the yield gap decomposition framework. 
- Stochastic frontiers analysis.
- So far only a Cobb Douglas production function.
- Limited to simpler models somewhat by the quality of the data.

## Data

- Main data source is the EGC-ISSER Ghana Panel Survey.
- One year of data covering 2009-10.
- We focus only on the major season as input use is low in the minor season.
- We find the quality of the data to be lower than the LSMS-ISA surveys. in general, and this results in a simpler analysis than Ethiopia.

## Summary statistics

\small
```{r echo=FALSE}
pander(readRDS(file.path(root, "Cache/sum_tab_res.rds")))
```
\normalsize

## Method

- Cobb Douglas production curve 
- We may experiment with more complicated production functions but initial results are not promising.
- Decomposition of the yield gap

## Frontier yield response model

\small
```{r echo=FALSE}
pander(readRDS(file.path(root, "Cache/CD_sf_res.rds")))
```
\normalsize

## Yield gaps by zone

\small 
```{r relative_YG}
ZonalYieldGap_l_sh <- ZonalYieldGap_l_sh %>%
  rename(FYG = EUYG)
names(ZonalYieldGap_l_sh) <- c("ZONE" ,"TEYG", "AYG", "EYG", "TYG", "YG")
knitr::kable(ZonalYieldGap_l_sh, caption = "Relative yield gap", digits = 0)
```
\normalsize

## National Yield gap decomposition

```{r Gap}
Fig_waterfall
```

## Economic yield assumptions

- Nitrogen capped at 200 kg/ha - link to nutrient gap project?
- Assets increased by 50% - is this reasonable?
- labour increased by 50% - is this reasonable?
- All farmers assumed to use herbicide
- All farmers assumed to use some form of mechanization in production process (tractor etc.)
- Other suggestions?
- Still to incorporate the GYGA variables in the analysis. This includes water-limited potential yield. Currently Pw = 15229 for all of Ghana!

## Next steps

- Add exogenous determinants of technical inefficiency. Policy implications.
- Calculate the marginal effects of exogenous determinants of technical inefficiency.
- Add the GYGA variables to control for growing conditions.
- Experiment with other functional forms - limited success so far.
- Endogeneity and heterogeneity? Data does not appear to be of sufficient quality to bear advanced techniques.

## Recommendations/Conclusions

- Smallholder farmers report very small yields compared to other countries e.g. Ethiopia, median is 466 kg/ha, consequently the frontier yield is also low 600 kg/ha.

- Largest yield gap is the technical yield gap, followed by the economic, allocative and technical efficiency yield gaps.

- Mean efficiency = 0.43

- Policies that close the allocative or economic yield gaps are very important. For example, subsidy programs which lower the cost of nitrogen will help to close the allocative yield gap.
- Determinants of inefficiency (still to come).
