---
title: "GHA Yield Gap Analysis"
author: "Michiel van Dijk & Tom Morley"
output:
  revealjs::revealjs_presentation:
    incremental: true
    theme: moon
    highlight: zenburn
    center: true
    transition: convex
    reveal_options:
      slideNumber: true


---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r get_data, echo = FALSE, message = FALSE}
# Get workspace with all info
dataPath <- "D:\\Data\\IPOP\\SurveyData\\"
load("D:\\Data\\Projects\\GHAYG\\Cache/FigTab.Rdata")

library(gridExtra)
library(dplyr)
source("D:/Data/Projects/GHAYG/Code/GHAmaps.R")
```

## Outstanding issues...solved

- Location of plots
    + GPS file linked
- Conversion of fertilizer units
    + Most units (e.g. bucket, maxi bag and American tin) converted with the suport of extension officers
    + Assumption made that all fertilizer is NPK (15/15/15)
    + Average use of fertilizer per hectare is in line with other information

## Location of plots
```{r plot_map}
grid.arrange(gmap1, gmap3, ncol=2)
```

## Summary statistics

```{r sum_stat,  results="asis"}
suppressMessages(library(stargazer))
stargazer(as.data.frame(dbsum),  type = "html", digits=2, header=FALSE, title = "Summary statistics")
```

## Frontier yield response model
```{r sfa}
knitr::kable(xtable, format = "html", padding = 0, digits = 2)
```



## Next steps
- Add exogenous variables, in particular growth defining factors
- Try different models
- Fix prices
- If possible, add variables that explain the yield gap in second stage analysis
- Perform yield gap analysis
